---
title: "Take Home Exercise 02:"
author: "Zhang Jinghan"
date-modified: "last-modified"
format: html
execute:
  eval: true
  echo: true
  warning: false
  freeze: true #防止重复render
---

# Oceanus Folk: Then-and-Now

# Data Preparation

## **Import R Packages**

For data cleaning, the following R packages will be used. They are **tidyverse**, **jsonlite**, **tidygraph** and **ggraph**.

```{r}
pacman::p_load(tidyverse, jsonlite,   
               SmartEDA, tidygraph, 
               ggraph,patchwork,DT,visNetwork,scales)
```

## **Import Data**

For the purpose of this exercise, *MC1_graph.json* file will be used. Before getting started, you should have the data set in the **data** sub-folder.

In the code chunk below, `fromJSON()` of **jsonlite** package is used to import *MC1_graph.json* file into R and save the output object

```{r}
kg <- fromJSON("data/MC1_graph.json")
```

## **Inspect Structure**

```{r}

str(kg, max.level = 1)  #show highest level的data structure
```

## Convert year-like character to integer

```{r}
# Convert year-like character columns into integer year
kg$nodes <- kg$nodes %>%
  mutate(
    release_date   = as.integer(release_date),
    written_date   = as.integer(written_date),
    notoriety_date = as.integer(notoriety_date)
  )

```

## **Extracting the edges and nodes tables**

Next, `as_tibble()` of **tibble** package package is used to extract the nodes and links tibble data frames from *kg* object into two separate tibble data frames called *nodes_tbl* and *edges_tbl* respectively.

```{r}
#Extract nodes and edges from kg
nodes_tbl <- as_tibble(kg$nodes)
edges_tbl <- as_tibble(kg$links)
```

# **Initial EDA**

**Frequency distribution of Node Type and Edge Type**

Uses ggplot2 functions to reveal the frequency distribution of *Node Type* field of *nodes_tbl*.

```{r}
#|code-fold: true
#|code-summary: "Click to View Code"
#|warning: false
#|message: false

# 图 1：Node Type 分布（按频数从高到低排序）
node_order <- nodes_tbl %>%
  count(`Node Type`) %>%
  arrange(n) %>%
  pull(`Node Type`)

p1 <- ggplot(data = nodes_tbl, aes(y = factor(`Node Type`, levels = node_order))) +
  geom_bar() +
  labs(title = "Node Type", y = NULL)

# 图 2：Edge Type 分布（按频数从高到低排序）
edge_order <- edges_tbl %>%
  count(`Edge Type`) %>%
  arrange(n) %>%
  pull(`Edge Type`)

p2 <- ggplot(data = edges_tbl, aes(y = factor(`Edge Type`, levels = edge_order))) +
  geom_bar() +
  labs(title = "Edge Type", y = NULL)

# 并排排列
p1 + p2

```

## Creating knowledge graph

### **Mapping from node id to row index**

Before we can go ahead to build the tidygraph object, it is important for us to ensures each id from the node list is mapped to the correct row number. This requirement can be achive by using the code chunk below.

```{r}
id_map <- tibble(id=nodes_tbl$id,
                 index = seq_len(
                   nrow(nodes_tbl)
                 ))
#处理后就没有id=0了
```

### **Map source and target IDs to row indices**

Next, we will map the source and the target IDs to row indices by using the code chunk below.

```{r}
edges_tbl <- edges_tbl %>%
  left_join(id_map, by = c("source" = "id")) %>%
  rename(from = index) %>%
  left_join(id_map, by = c("target" = "id")) %>%
  rename(to = index)
```

### **Filter out any unmatched (invalid) edges**

Lastly, the code chunk below will be used to exclude the unmatch edges.

```{r}
edges_tbl <- edges_tbl%>%
  filter(!is.na(from), !is.na(to))
```

### **Creating tidygraph**

Lastly, `tbl_graph()` is used to create tidygraph's graph object by using the code chunk below.

```{r}
graph <- tbl_graph(nodes = nodes_tbl, 
                   edges = edges_tbl, 
                   directed = kg$directed)
```

You might want to confirm the output object is indeed in tidygraph format by using the code chunk below.

```{r}
class(graph)
```

### **Visualising the knowledge graph**

In this section, we will use ggraph's functions to visualise and analyse the graph object.Visualising the knowledge graph.

Several of the ggraph layouts involve randomisation. In order to ensure reproducibility, it is necessary to set the seed value before plotting by using the code chunk below.

```{r}
set.seed(1234)
```

```{r}
#| eval: false
#| freeze: true
ggraph(graph, layout = "fr") +      
  geom_edge_link(alpha = 0.3, 
                 colour = "gray") +
  geom_node_point(aes(color = `Node Type`), 
                  size = 1) +
  geom_node_text(aes(label = name), 
                 repel = TRUE, 
                 size = 2.5) +
  theme_void()

```

------------------------------------------------------------------------

# Q2: The Spread of Oceanus Folk

## EDA

### Type of Nodes and Edges

```{r}
#| code-fold: true
# 图 1：Node Type 分布（按频数从高到低排序）
node_order <- nodes_tbl %>%
  count(`Node Type`) %>%
  arrange(n) %>%
  pull(`Node Type`)

p1 <- ggplot(data = nodes_tbl, aes(y = factor(`Node Type`, levels = node_order))) +
  geom_bar() +
  labs(title = "Node Type", y = NULL)

# 图 2：Edge Type 分布（按频数从高到低排序）
edge_order <- edges_tbl %>%
  count(`Edge Type`) %>%
  arrange(n) %>%
  pull(`Edge Type`)

p2 <- ggplot(data = edges_tbl, aes(y = factor(`Edge Type`, levels = edge_order))) +
  geom_bar() +
  labs(title = "Edge Type", y = NULL)

# 并排排列
p1 + p2
```

|   | Edge Type | From Node Type (A) | To Node Type (B) | Flow of Influence |
|----|----|----|----|----|
| Produce | **DistributedBy** | song/album | record label | / |
| Produce | **RecordedBy** | song/album | record label | / |
| Contribute/Create | **ComposerOf** | person | song/album | A to B |
| Contribute/Create | **ProducerOf** | person | song/album | A to B |
| Contribute/Create | **LyricistOf** | person | song/album | A to B |
| Contribute/Create | **PerformerOf** | person | song/album | A to B |
| Belong to | **MemberOf** | person | music group | / |
| Influence | **InStyleOf** | song/album | song/album | B to A |
| Influence | **InterpolatesFrom** | song/album | song/album | B to A |
| Influence | **LyricalReferenceTo** | song/album | song/album | B to A |
| Influence | **CoverOf** | song/album | song/album | B to A |
| Influence | **DirectlySamples** | song/album | song/album | B to A |

: The table above summarise all the node types and edge types in the knowledge graph, highlighting the direction and meaning of influence between them.

### Type of Genres

There are 26 genres in total with dream pop having most songs and albums. Oceanus Folk ranks 5th with 70 albums and 235 songs in total.

```{r}
#| code-fold: true
#| warning: false
#| message: false
#| freeze: true
# Count the number of Songs and Albums in each genre
genre_summary <- nodes_tbl %>%
  filter(`Node Type` %in% c("Song", "Album")) %>%
  count(genre, `Node Type`) %>%
  pivot_wider(names_from = `Node Type`, values_from = n, values_fill = 0) %>%
  arrange(desc(Song + Album))


print(genre_summary)
```

```{r}
#| code-fold: true
#| warning: false
#| message: false

# 1. Sort by song count
genre_order <- genre_summary %>%
  arrange(desc(Song)) %>%
  pull(genre)

# 2. transform to long table
genre_long <- genre_summary %>%
  pivot_longer(cols = c(Song, Album), names_to = "Type", values_to = "Count") %>%
  mutate(genre = factor(genre, levels = genre_order))

# 3. plot the chart
ggplot(genre_long, aes(x = genre, y = Count, fill = Type)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +
  #geom_text(aes(label = Count),
            #position = position_dodge(width = 0.9),
            #hjust = 1.05, color = "black", size = 3.2) +
  labs(
    title = "Number of Songs and Albums by Genre (Sorted by Song Count)",
    x = "Genre",
    y = "Count"
  ) +
  theme_minimal() +
  coord_flip()
```

### Songs

```{r}
#| code-fold: true
#| warning: false
#| message: false
# Step 1: Filter all Song nodes
songs_tbl <- nodes_tbl %>%
  filter(`Node Type` == "Song")

# Step 2: Count total number of songs
total_songs <- nrow(songs_tbl)

# Step 3: Count number of notable songs
notable_songs <- songs_tbl %>%
  filter(notable == TRUE) %>%
  nrow()

# Step 4: Calculate proportion
notable_ratio <- round(notable_songs / total_songs, 3)

# Output summary (safe version, without emojis)
tibble(
  Total_Songs = total_songs,
  Notable_Songs = notable_songs,
  Proportion_Notable = notable_ratio
)
```

```{r}
#| code-fold: true
#| warning: false
#| message: false
# Step 1: Filter song nodes with valid release date
songs_by_year <- nodes_tbl %>%
  filter(`Node Type` == "Song", !is.na(release_date)) %>%
  count(release_date, name = "count")

# Step 2: Plot chart using all defaults
ggplot(songs_by_year, aes(x = release_date, y = count)) +
  geom_col() +
  labs(
    title = "Number of Songs Released by Year (All Genre)",
    x = "Release Year",
    y = "Number of Songs"
  ) +
  theme_minimal()

```

```{r}
#| code-fold: true
#| warning: false
#| message: false
# Step 1: Filter Oceanus Folk Songs and Albums with release date
oceanus_by_year <- nodes_tbl %>%
  filter(genre == "Oceanus Folk",
         `Node Type` %in% c("Song", "Album"),
         !is.na(release_date)) %>%
  count(release_date, `Node Type`, name = "count")

# Step 2: Plot chart with dodged bars
ggplot(oceanus_by_year, aes(x = release_date, y = count, fill = `Node Type`)) +
  geom_col(position = "dodge", width = 0.7) +
  labs(
    title = "Number of Oceanus Folk Songs and Albums by Year",
    x = "Release Year",
    y = "Count"
  ) +
  scale_fill_manual(values = c("Song" = "#2f4b7c", "Album" = "#b2defd")) +
  theme_minimal()

```

```{r}
#| code-fold: true
#| warning: false
#| message: false
#| freeze: true
#Step 1: Prepare data — total songs and Oceanus Folk songs by year
songs_by_year <- nodes_tbl %>%
  filter(`Node Type` == "Song", !is.na(release_date)) %>%
  mutate(is_oceanus = ifelse(genre == "Oceanus Folk", "Oceanus Folk", "All Songs")) %>%
  count(release_date, is_oceanus, name = "count")

# Step 2: Plot line chart comparing the two series
ggplot(songs_by_year, aes(x = release_date, y = count, color = is_oceanus)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  geom_text(aes(label = count), vjust = -0.6, size = 3) +
  scale_color_manual(values = c("All Songs" = "#525252", "Oceanus Folk" = "#2f4b7c")) +
  labs(
    title = "Number of Songs by Year: Oceanus Folk vs All Songs",
    x = "Release Year",
    y = "Number of Songs",
    color = NULL
  ) +
  theme_minimal()

```

## 2a. Was this influence intermittent or did it have a gradual rise?

::: callout-note
### Assumption

We assume that the **influence of Oceanus Folk** increases as the **number of works released** in the genre grows. A higher release volume suggests broader artist adoption and greater genre presence in the music network.
:::

### **2a.1 Number of Oceanus Folk Release**

This visualization combines a yearly line chart and a 5-year binned heatmap to analyze the release trend of Oceanus Folk works.

#### **Step 1: Prepare yearly release data**

The dataset is filtered to include only `Song` and `Album` nodes where the genre is "Oceanus Folk". The number of releases per year is counted and reshaped using `pivot_wider` to separate counts for Songs and Albums. A new column `total` is created to represent the yearly sum of both types.

#### **Step 2: Generate 5-year bins for heatmap**

Years are grouped into 5-year bins using `floor(release_date / 5) * 5`.\
For each bin, the total number of releases, as well as breakdowns for albums and songs, are aggregated. An additional column `change` calculates the absolute difference in total releases from the previous bin. A hover text is prepared for display in the interactive plot.

#### **Step 3: Construct the static ggplot**

The plot contains two layers: - A **heatmap** (`geom_tile`) aligned below the x-axis, using fill color to represent the total number of releases in each 5-year bin. - A **line chart** (`geom_line`) showing the trend of total yearly releases, with **points** (`geom_point`) that reveal detailed hover information for each year, including totals and breakdowns.

#### **Step 4: Convert to interactive plot**

The static plot is transformed into an interactive visualization using `ggplotly()`. The tooltip is set to use the custom `text` field to display detailed year or period-specific data when users hover over the plot.

This combined approach allows users to explore both fine-grained yearly trends and broader multi-year patterns in the adoption of Oceanus Folk genre.

```{r}
#| code-fold: true
#| warning: false
#| message: false
#| freeze: true
library(plotly)

# Step 1: Count number of Oceanus Folk songs and albums by year
oceanus_by_year <- nodes_tbl %>%
  filter(genre == "Oceanus Folk",
         `Node Type` %in% c("Song", "Album"),
         !is.na(release_date)) %>%
  count(release_date, `Node Type`, name = "count") %>%
  pivot_wider(names_from = `Node Type`, values_from = count, values_fill = 0) %>%
  mutate(total = Song + Album)

# Step 2: Create 5-year bins and compute heatmap values and absolute change
oceanus_by_year <- oceanus_by_year %>%
  mutate(year_bin = floor(release_date / 5) * 5)

heatmap_df <- oceanus_by_year %>%
  group_by(year_bin) %>%
  summarise(
    bin_total = sum(total),
    Album = sum(Album),
    Song = sum(Song),
    .groups = "drop"
  ) %>%
  arrange(year_bin) %>%
  mutate(
    change = bin_total - lag(bin_total),
    hover_text = paste0("Period: ", year_bin, "–", year_bin + 4,
                        "<br>Total: ", bin_total,
                        "<br>Change: ", ifelse(is.na(change), "NA", 
                                        ifelse(change > 0, paste0("+", change), change)))
  )

# Step 3: Build static ggplot with line and heatmap
p <- ggplot() +
  # Add heatmap tiles below the line
  geom_tile(data = heatmap_df,
            aes(x = year_bin + 2.5, y = -5, fill = bin_total, 
                text = hover_text),
            width = 5, height = 3) +
  # Add line for total releases
  geom_line(data = oceanus_by_year,
            aes(x = release_date, y = total, group = 1), 
            color = "#2f4b7c", linewidth = 1) +
  # Add points with interactive hover labels
  geom_point(data = oceanus_by_year,
             aes(x = release_date, y = total,
                 text = paste0("Year: ", release_date,
                               "<br>Total: ", total,
                               "<br>Album: ", Album,
                               "<br>Song: ", Song)),
             color = "#2f4b7c", size = 2) +
  scale_fill_gradient(low = "#a8dadc", high = "#2f4b7c") +
  labs(
    title = "Oceanus Folk Releases by Year with 5-Year Interval Heatmap",
    x = "Release Year",
    y = "Total Count",
    fill = "5-Year Total"
  ) +
  theme_minimal()

# Step 4: Convert to interactive plotly object
ggplotly(p, tooltip = "text")

```

The line chart above plots the number of release in Oceanus folk genre by year while the heat map bar at bottom is based on 5 year binned number of release in Oceanus folk genre. Darker the colour is means greater number of songs or album published during the binned period.

::: callout-note
## Insights:

We assume that higher volume of productions in a genre reflects the influence and spread of genre because it suggests greater adoption by artists, indicating that the genre resonates with creators and is being actively used as a mode of expression.

Oceanus Folk's rise to prominence appears to be a two-phase process: it began with a gradual buildup from the 1990s to the late 2000s, followed by a sudden surge in releases from 2010 to 2014. After that there was a second surge around 2020 to 2024 and peaked at the same period.

**This pattern suggests that the genre first accumulated a foundational base of creators and early adopters, then experienced a rapid acceleration in influence and production, likely driven by cultural resonance, landmark works, or wider adoption across communities.**
:::

## 2b. What genres and top artists have been most influenced by Oceanus Folk

::: callout-note
### Assumption

We assume that **the more connections a genre or artist has to Oceanus Folk–influenced works**, the greater the **interaction** and **influence**.

A higher number of such links suggests that Oceanus Folk has been actively referenced, adopted, or reinterpreted by other creators.

### Definition of Influence between Genres via Song/Album

Influence in this context is defined based on specific edge types that indicate a stylistic or content-based relationship.\
The following edge types are used to capture different modes of influence:

-   `InStyleOf`: The influenced work adopts the style of an Oceanus Folk work
-   `CoverOf`: A direct cover or reinterpretation
-   `InterpolatesFrom`: Elements of the original are restructured or embedded
-   `DirectlySamples`: A portion of the Oceanus Folk work is directly used
-   `LyricalReferenceTo`: The lyrics reference Oceanus Folk content

These relationships are traced from Oceanus Folk songs and albums to other works, forming the basis for calculating influence metrics on genres and artists.
:::

### 2b.1 Transform Genre to Virtual Node

Currently, genre is an attribute of song/album nodes in the dataset, we need to transform it to a virtual node.

To construct the influence network of the Oceanus Folk genre, we begin by enriching the original edge list (`edges_tbl`) with metadata about node types and genres. Songs and albums classified under the Oceanus Folk genre are identified as source nodes, and influence-type edges (e.g., *InStyleOf*, *CoverOf*, *InterpolatesFrom*) originating from these works are selected.

Next, we define a central "Oceanus Folk" genre node to act as a hub, and extract metadata about works influenced by Oceanus Folk. These works are represented as nodes, and directed edges from the genre node to the works are constructed. Additional genre nodes are created for the influenced works, and a secondary set of edges links these works to their respective genres.

Node sizes are scaled based on the number of incoming connections (in-degree), representing how many times a work or genre has been referenced. The Oceanus Folk genre node is manually assigned a larger size to visually emphasize its centrality.

Finally, we assign visual properties to the edges. Each edge is colored according to its influence type, and tooltips displaying the edge type are included using the `title` field. To maintain a clean visual layout, the default `label` field is removed to avoid persistent on-edge text. This ensures that edge types are only revealed upon hovering, thereby reducing visual clutter.

```{r}
#| code-fold: true
#| warning: false
#| message: false
#| freeze: true
library(dplyr)
library(scales)

# Step 1: Enrich edge metadata with source and target node types and genres
edge_graphdata <- edges_tbl %>%
  left_join(nodes_tbl %>% select(id, genre_from = genre),
            by = c("from" = "id")) %>%
  left_join(nodes_tbl %>% select(id, type_from = `Node Type`),
            by = c("from" = "id")) %>%
  left_join(nodes_tbl %>% select(id, type_to = `Node Type`),
            by = c("to" = "id"))

# Step 2: Identify all Oceanus Folk songs and albums
of_publications <- nodes_tbl %>%
  filter(`Node Type` %in% c("Song", "Album"), genre == "Oceanus Folk") %>%
  pull(id)

# Step 3: Select influence-type edges originating from Oceanus Folk works
of_outgoing_edges <- edge_graphdata %>%
  filter(from %in% of_publications,
         `Edge Type` %in% c("InStyleOf", "CoverOf", "InterpolatesFrom",
                            "DirectlySamples", "LyricalReferenceTo"))

# Step 4: Collect metadata of target works influenced by Oceanus Folk
target_metadata <- of_outgoing_edges %>%
  left_join(nodes_tbl %>%
              select(id, name_target = name, genre_target = genre, type_target = `Node Type`),
            by = c("to" = "id")) %>%
  distinct(to, name_target, genre_target, type_target)

# Step 5: Create a virtual node representing the Oceanus Folk genre
node_of <- tibble(
  id    = "Oceanus Folk",
  label = "Oceanus Folk",
  group = "Genre",
  title = "Genre node for Oceanus Folk<br>Aggregated influence hub"
)

# Step 6: Create nodes for the influenced works (only Song and Album types)
target_graphdata <- target_metadata %>%
  filter(type_target %in% c("Song", "Album")) %>%
  transmute(
    id    = as.character(to),
    label = name_target,
    group = type_target,
    title = paste0("Genre: ", genre_target)
  )

# Step 7: Create influence edges from Oceanus Folk to its influenced works
edge_of_to_target <- of_outgoing_edges %>%
  left_join(nodes_tbl %>% select(id, `Node Type`), by = c("to" = "id")) %>%
  filter(`Node Type` %in% c("Song", "Album")) %>%
  transmute(
    from  = "Oceanus Folk",
    to    = as.character(to),
    title = paste0("Edge Type: ", `Edge Type`)
  )

# Step 8A: Create nodes for all other genre types (excluding Oceanus Folk)
other_genres <- target_metadata %>%
  filter(!is.na(genre_target), genre_target != "Oceanus Folk") %>%
  distinct(genre_target) %>%
  transmute(
    id = genre_target,
    label = genre_target,
    group = "Genre",
    title = paste0("Genre node: ", genre_target)
  )

# Step 8B: Create edges from works to their corresponding genres
edge_work_to_genre <- target_metadata %>%
  filter(!is.na(genre_target), 
         genre_target != "Oceanus Folk",
         type_target %in% c("Song", "Album")) %>%
  transmute(
    from = as.character(to),
    to   = genre_target,
    title = "Belongs to genre"
  )

# Step 9: Assemble the full node and edge tables
node_graphdata <- bind_rows(node_of, target_graphdata, other_genres) %>%
  distinct(id, .keep_all = TRUE)

edge_graphdata_final <- bind_rows(edge_of_to_target, edge_work_to_genre) %>%
  distinct(from, to, .keep_all = TRUE)

# Step 10: Compute in-degree (number of incoming edges per node)
target_meta <- edge_graphdata_final %>%
  count(to, name = "n_incoming")

# Step 11: Assign node sizes based on in-degree
node_graphdata <- node_graphdata %>%
  left_join(target_meta, by = c("id" = "to")) %>%
  mutate(size = case_when(
    id == "Oceanus Folk" ~ 60,
    TRUE ~ rescale(n_incoming, to = c(20, 50), na.rm = TRUE))
  ) %>%
  replace_na(list(size = 20))

# Step 12: Finalize edge visualization with type, color, and hover tooltip
edges_vis <- edge_graphdata_final %>%
  mutate(
    label = case_when(
      str_detect(title, regex("InStyleOf", ignore_case = TRUE)) ~ "InStyleOf",
      str_detect(title, regex("CoverOf", ignore_case = TRUE)) ~ "CoverOf",
      str_detect(title, regex("InterpolatesFrom", ignore_case = TRUE)) ~ "InterpolatesFrom",
      str_detect(title, regex("DirectlySamples", ignore_case = TRUE)) ~ "DirectlySamples",
      str_detect(title, regex("LyricalReferenceTo", ignore_case = TRUE)) ~ "LyricalReferenceTo",
      str_detect(title, regex("Belongs to genre", ignore_case = TRUE)) ~ "BelongsToGenre",
      TRUE ~ "Other"
    ),
    color = case_when(
      label == "InStyleOf" ~ "blue",
      label == "CoverOf" ~ "red",
      label == "InterpolatesFrom" ~ "purple",
      label == "DirectlySamples" ~ "orange",
      label == "LyricalReferenceTo" ~ "green",
      label == "BelongsToGenre" ~ "gray",
      TRUE ~ "black"
    ),
    width = 2,
    title = paste0("Edge Type: ", label)
  )


```

### **2b.2 Plotting Oceanus Folk Influence Network**

We plot a network with *Oceanus Folk* at the center, connected to the works it has influenced and the genres those works belong to — forming an **Oceanus Folk → Influenced Work → Genre** structure. This structure helps us identify which genres have been most influenced by Oceanus Folk.

1.  **Standardize node types**\
    Nodes originally labeled as "Song" or "Album" are reclassified under a unified group "Work" for consistent styling and group-based coloring.

2.  **Extract and format genre information**

    -   For work nodes, the genre is extracted from the node title and stored in a separate `display_genre` field.
    -   For genre nodes, the label "Genre node" is assigned.

3.  **Customize tooltips**\
    HTML-formatted tooltips (`title`) are defined for each node. These tooltips include the node’s name, type, and genre (if available), and appear when the user hovers over the node.

4.  **Define node colors**\
    Colors are assigned based on node type:

    -   Genre nodes → `skyblue`
    -   Work nodes (songs/albums) → `lightgreen`
    -   Others → `gray`

5.  **Control label visibility**\
    Only genre nodes are assigned visible labels (with a larger font size). Work nodes remain unlabeled to reduce clutter.

6.  **Render the interactive network**\
    Using `visNetwork()`, the node and edge data is rendered with the following key features:

    -   **Highlighting and interactivity:** Users can hover over or select nodes to explore their connections.
    -   **Directed arrows:** All edges are directional (`arrows = "to"`), indicating influence or genre linkage.
    -   **Physics engine:** The *forceAtlas2Based* physics solver is used for layout, with gravitational tuning and stabilization for a clean layout.
    -   **Legend configuration:**
        -   Node group legend is auto-generated using `visGroups()`.
        -   Edge type legend is manually added via `addEdges` in `visLegend()`, matching edge colors to their semantic meanings (e.g., "CoverOf" = red).

```{r}
#| code-fold: true
#| warning: false
#| message: false
#| freeze: true

nodes_vis <- node_graphdata %>%
  mutate(
    # Standardize node types
    group = case_when(
      group %in% c("Song", "Album") ~ "Work",
      TRUE ~ group
    ),
    # Extract genre information
    display_genre = case_when(
      group == "Work" ~ str_extract(title, "(?<=Genre: ).*"),
      group == "Genre" ~ "Genre node",
      TRUE ~ NA_character_
    ),
    # Set HTML-formatted tooltips
    title = paste0(
      "<b>", label, "</b><br>",
      "Type: ", group, "<br>",
      ifelse(!is.na(display_genre), paste0("Genre: ", display_genre, "<br>"), "")
    ),
    # Assign node colors
    color = case_when(
      group == "Genre"  ~ "skyblue",
      group == "Work"   ~ "lightgreen",
      TRUE              ~ "gray"
    ),
    # Only genre nodes display labels
    font_size = ifelse(group == "Genre", 20, 0)
  ) %>%
  mutate(font = map(font_size, ~ list(size = .x)))

visNetwork(nodes = nodes_vis, edges = edges_vis, 
           width = "100%", height = "700px") %>%
  visNodes(
    borderWidth = 2,
    borderWidthSelected = 4,
    shadow = TRUE,
    color = list(
      border = "black",
      highlight = list(border = "black", background = "orange")
    )
  ) %>%
  visEdges(arrows = "to", smooth = FALSE) %>%
  visOptions(
    highlightNearest = TRUE,
    nodesIdSelection = list(enabled = TRUE, useLabels = TRUE)
  ) %>%
  visInteraction(zoomView = TRUE) %>%
  visLayout(randomSeed = 1234) %>%
  visGroups(groupname = "Genre", color = "skyblue") %>%
  visGroups(groupname = "Work", color = "lightgreen") %>%
  visLegend(
    useGroups = TRUE,
    position = "right",
    main = "Legend",
    addEdges = data.frame(
      label = c("InStyleOf", "CoverOf", "InterpolatesFrom", 
                "DirectlySamples", "LyricalReferenceTo", "BelongsToGenre"),
      color = c("blue", "red", "purple", "orange", "green", "gray")
    )
  ) %>%
  visPhysics(
    solver = "forceAtlas2Based",
    forceAtlas2Based = list(gravitationalConstant = -50),
    stabilization = list(enabled = TRUE, iterations = 100)
  )


```

**Network Structure**

This visualization clearly communicates the influence pathway of *Oceanus Folk* as a genre. The network is structured in a two-hop flow: \[**Genre → Work → Genre**\]

-   The central blue node labeled **Oceanus Folk** serves as the origin point of influence.

-   It connects outward to multiple **light green nodes**, which represent the **musical works** influenced by Oceanus Folk.

-   Each of these works is further linked to a **sky blue node**, indicating the **genre** the work belongs to.

Each colored arrow denotes a specific type of influence, such as *InStyleOf*, *CoverOf*, or *DirectlySamples*, making it possible to distinguish how the influence was transmitted.

This structure helps us identify which genres have been most influenced by Oceanus Folk through intermediary works — notably genres like **Indie Folk**, **Synthwave**, and **Dream Pop**, which appear as large blue nodes due to their high number of incoming connections.

::: callout-note
#### Insight

**Indie Folk** is the genre which receives most influence in degree from Oceanus Folk reflected as the biggest node except Oceanus Folk in the network, Indicating high centrality and cumulative influence.

**Deam Pop** ranks second, absorbed notable influence from Oceanus Folk through works *In Style Of* or *Interpolates From* Oceanus Folk.

**Synthwave** comes third but shows a strong connection via Directly Samples relationships.
:::

### 2b.3 Top Influenced Genres: Absolute and Relative Analysis

This part aims to offers insights into which musical styles most absorbed Oceanus Folk's influence, both in absolute and proportional terms.

We first counted how many influenced works were ultimately labeled under each genre using `BelongsToGenre` edges. This provides an **absolute measure** of genre influence. Next, to account for the varying total output sizes of different genres, we joined this with the **total number of works per genre**, allowing us to compute a **relative influence ratio**: the share of influenced works out of total genre works.

```{r}
top_genre_rank <- edges_vis %>%
  filter(label == "BelongsToGenre") %>%
  count(to, name = "influenced_work_count") %>%
  arrange(desc(influenced_work_count)) %>%
  left_join(nodes_vis %>% select(id, label), by = c("to" = "id")) %>%
  rename(genre = label)

#top_genre_rank %>% slice_head(n = 5)

genre_total_count <- nodes_tbl %>%
  filter(`Node Type` %in% c("Song", "Album")) %>%
  count(genre, name = "total_works")

genre_summary <- top_genre_rank %>%
  left_join(genre_total_count, by = c("genre" = "genre")) %>%
  mutate(ratio = influenced_work_count / total_works)


genre_summary %>%
  arrange(desc(ratio)) %>%
  slice_head(n = 5)
```

Two complementary visualizations are presented:

```{r}
library(ggplot2)
library(patchwork)

# Plot 1: Absolute Influenced Works
p1 <- ggplot(genre_summary, aes(x = reorder(genre, influenced_work_count), 
                                y = influenced_work_count)) +
  geom_col(fill = "skyblue") +
  coord_flip() +
  labs(
    title = "Top Influenced Genres by Oceanus Folk (Absolute Count)",
    x = "Genre",
    y = "Number of Influenced Works"
  ) +
  theme_minimal(base_size = 14) +
  theme(plot.title = element_text(lineheight = 1.1, hjust = 0.5))

# Plot 2: Relative Influenced Share (Lollipop)
p2 <- ggplot(genre_summary, aes(x = ratio, y = reorder(genre, ratio))) +
  geom_segment(aes(x = 0, xend = ratio, yend = genre), color = "gray") +
  geom_point(color = "steelblue", size = 5) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    title = "Top Influenced Genres by Oceanus Folk (Relative Share)",
    x = "Influenced Share of Works",
    y = "Genre"
  ) +
  theme_minimal(base_size = 14) +
  theme(plot.title = element_text(lineheight = 1.1, hjust = 0.5))

# Combine vertically
p1 / p2

```

**Bar Chart (Top Panel):** Shows the absolute number of works in each genre influenced by Oceanus Folk.

-   *Indie Folk* leads with 7 works, followed by *Dream Pop* (4), *Synthwave* (3), *Celtic Folk* (2), and *Acoustic Folk* (1). These counts reflect the **breadth of Oceanus Folk's influence** across stylistic boundaries.

**Lollipop Plot (Bottom Panel):** Presents the relative influence — the proportion of each genre’s total works that were influenced.

-   Though *Celtic Folk* has only 2 influenced works, it ranks high in **relative share** due to a small base of total works, highlighting a **concentrated influence**. This view helps distinguish niche genres with deep stylistic borrowings from those with broad but diffuse connections.

Based on the number of works influenced by Oceanus Folk and subsequently categorized under different genres, **Indie Folk** ranks first with 7 influenced works, indicating a strong stylistic connection. **Dream Pop** follows with 4 works, showing notable indirect influence. **Synthwave** takes the third position with 3 works, some of which may reflect direct sampling relationships. **Celtic Folk** and **Acoustic Folk** complete the top five, with 2 and 1 influenced works respectively.

Total number of works each genre have

```{r}

```

```{r}


```

```         
```

```{r}
library(ggplot2)
library(scales)

# Calculate scale factor
max_count <- max(genre_summary$influenced_work_count)
max_ratio <- max(genre_summary$ratio)
scale_factor <- max_count / max_ratio

# Add scaled ratio for plotting
plot_data <- genre_summary %>%
  mutate(scaled_ratio = ratio * scale_factor)

ggplot(plot_data, aes(x = reorder(genre, influenced_work_count))) +
  # Bar: absolute influenced count
  geom_col(aes(y = influenced_work_count, fill = "Influenced Count"), width = 0.6) +

  # Line + point: relative ratio (scaled)
  geom_line(aes(y = scaled_ratio, color = "Influence Ratio", group = 1), size = 1.2) +
  geom_point(aes(y = scaled_ratio, color = "Influence Ratio"), size = 3) +

  # Axes
  scale_y_continuous(
    name = "Number of Influenced Works",
    sec.axis = sec_axis(~ . / scale_factor,
                        name = "Influenced Share",
                        labels = percent_format(accuracy = 1))
  ) +
  labs(
    title = "Oceanus Folk's Influence on Genres (Absolute vs. Relative)",
    x = "Genre",
    fill = "",
    color = ""
  ) +
  scale_fill_manual(values = c("Influenced Count" = "skyblue")) +
  scale_color_manual(values = c("Influence Ratio" = "steelblue")) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "top",
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),  # genre竖着显示
    axis.title.y.right = element_text(color = "steelblue"),
    axis.text.y.right = element_text(color = "steelblue"),
    axis.line.y.right = element_line(color = "steelblue"),
    axis.ticks.y.right = element_line(color = "steelblue")
  )

```

```{r}
library(ggplot2)
library(ggrepel)
library(dplyr)

# Step 1: 计算中位数或均匀分位数作为阈值
x_median <- median(genre_summary$influenced_work_count)
y_median <- median(genre_summary$ratio)

# Step 2: 添加象限标签
genre_quadrant <- genre_summary %>%
  mutate(
    quadrant = case_when(
      influenced_work_count >= x_median & ratio >= y_median ~ "I: High-High",
      influenced_work_count <  x_median & ratio >= y_median ~ "II: Low-High",
      influenced_work_count <  x_median & ratio <  y_median ~ "III: Low-Low",
      influenced_work_count >= x_median & ratio <  y_median ~ "IV: High-Low"
    )
  )

# Step 3: 绘图
ggplot(genre_quadrant, aes(x = influenced_work_count, y = ratio, color = quadrant)) +
  geom_point(size = 4) +
  geom_vline(xintercept = x_median, linetype = "dashed", color = "gray50") +
  geom_hline(yintercept = y_median, linetype = "dashed", color = "gray50") +
  geom_text_repel(aes(label = genre), size = 4) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(
    x = "Number of Influenced Works",
    y = "Share of Works Influenced",
    title = "Oceanus Folk's Influence by Genre",
    subtitle = "Divided into Four Quadrants by Median Thresholds",
    color = "Quadrant"
  ) +
  theme_minimal(base_size = 14)

```

------------------------------------------------------------------------

## Top Artists who are most influenced by Oceanus Folk

```{r}
# Step 0: 筛选有contribute to notable works的top artist
# 1. 找出所有notable works
notable_works <- nodes_tbl %>%
  filter(`Node Type` %in% c("Song", "Album"), notable == TRUE) %>%
  pull(id)

# 2. 找出直接contribute的artist
contrib_types <- c("PerformerOf", "ComposerOf", "ProducerOf", "LyricistOf")
direct_contributors <- edges_tbl %>%
  filter(`Edge Type` %in% contrib_types, to %in% notable_works) %>%
  pull(from)

# 3. 可选：group成员（如果有group对notable work有贡献）
group_ids <- nodes_tbl %>%
  filter(`Node Type` == "Music Group") %>%
  pull(id)
group_contributors <- edges_tbl %>%
  filter(`Edge Type` %in% contrib_types, from %in% group_ids, to %in% notable_works) %>%
  pull(from)
group_members <- edges_tbl %>%
  filter(`Edge Type` == "MemberOf", to %in% group_contributors) %>%
  pull(from)

# 4. 合并所有top artist id
# 注意：有些id可能不是Person类型，后续分析时需与Person节点交集
all_top_artist_ids <- unique(c(direct_contributors, group_members))

# 5. 只保留Person类型的top artist id
person_ids <- nodes_tbl %>%
  filter(`Node Type` == "Person") %>%
  pull(id)
top_artist_ids <- intersect(all_top_artist_ids, person_ids)
```

```{r}
# 📦 Load libraries
library(dplyr)
library(purrr)
library(stringr)
library(visNetwork)
library(scales)

# 📌 Step 1: enrich edge metadata
edge_graphdata <- edges_tbl %>%
  left_join(nodes_tbl %>% select(id, genre_from = genre),
            by = c("from" = "id")) %>%
  left_join(nodes_tbl %>% select(id, type_from = `Node Type`),
            by = c("from" = "id")) %>%
  left_join(nodes_tbl %>% select(id, type_to = `Node Type`),
            by = c("to" = "id"))

# 📌 Step 2: get all Oceanus Folk publications (Song or Album)
of_publications <- nodes_tbl %>%
  filter(`Node Type` %in% c("Song", "Album"), genre == "Oceanus Folk") %>%
  pull(id)

# 📌 Step 3: get edges representing influence from OF works to others
influence_edge_types <- c("InStyleOf", "CoverOf", "InterpolatesFrom",
                          "DirectlySamples", "LyricalReferenceTo")

of_outgoing_edges <- edge_graphdata %>%
  filter(from %in% of_publications, `Edge Type` %in% influence_edge_types)

# 📌 Step 4: get metadata for influenced works
influenced_work_metadata <- of_outgoing_edges %>%
  left_join(nodes_tbl %>%
              select(id, name_work = name, type_work = `Node Type`),
            by = c("to" = "id")) %>%
  distinct(to, name_work, type_work)

# 📌 Step 5: get people who are linked to这些works ← 🔁 REVERSED DIRECTION HERE!
person_link_types <- c("PerformerOf", "ComposerOf", "ProducerOf", "LyricistOf")

work_to_person_edges <- edge_graphdata %>%
  filter(to %in% influenced_work_metadata$to,
         `Edge Type` %in% person_link_types) %>%
  left_join(nodes_tbl %>%
              filter(`Node Type` == "Person") %>%
              select(id, name_person = name),
            by = c("from" = "id")) %>%
  transmute(
    from = as.character(to),        # from: song
    to   = as.character(from),      # to: person
    title = paste0("Role: ", `Edge Type`),
    name_person
  ) %>%
  filter(!is.na(name_person), to %in% top_artist_ids) %>%  # 只保留top artist
  distinct(from, to, title, name_person)

# 📌 Step 6: define nodes
node_of <- tibble(
  id = "Oceanus Folk",
  label = "Oceanus Folk",
  group = "Genre",
  title = "Genre node: Oceanus Folk"
)

node_work <- nodes_tbl %>%
  filter(id %in% influenced_work_metadata$to) %>%
  transmute(
    id = as.character(id),
    label = name,
    group = `Node Type`,  # original type: Song or Album
    title = paste0("Influenced Work: ", name)
  )

node_person <- work_to_person_edges %>%
  distinct(to, name_person) %>%
  transmute(
    id = to,
    label = name_person,
    group = "Person",
    title = paste0("Artist: ", name_person)
  )

# 📌 Step 7: define edges
edge_of_to_work <- of_outgoing_edges %>%
  transmute(
    from = "Oceanus Folk",
    to   = as.character(to),
    title = paste0("Influence: ", `Edge Type`)
  )

edge_work_to_person <- work_to_person_edges %>%
  select(from, to, title)



  
# 📌 Step 8: assemble full node table with only connected nodes

# 合并所有节点候选：Oceanus Folk + 被影响的作品 + 参与的人
node_graphdata_full <- bind_rows(node_of, node_work, node_person) %>%
  distinct(id, .keep_all = TRUE)

# 获取所有出现在边中的节点 ID（from 和 to 都算）
used_node_ids <- unique(c(edge_of_to_work$from, edge_of_to_work$to,
                          edge_work_to_person$from, edge_work_to_person$to))

# 仅保留参与连接的节点
node_graphdata <- node_graphdata_full %>%
  filter(id %in% used_node_ids) %>%
  distinct(id, .keep_all = TRUE)
  


# 📌 Step 9: compute in-degree (number of incoming edges)
target_meta <- edge_graphdata_final %>%
  count(to, name = "n_incoming")

node_graphdata <- node_graphdata %>%
  mutate(id = as.character(id)) %>%
  left_join(target_meta, by = c("id" = "to")) %>%
  mutate(size = case_when(
    id == "Oceanus Folk" ~ 60,
    TRUE ~ rescale(n_incoming, to = c(10, 50), na.rm = TRUE))
  ))

# 📌 Step 10: format for visNetwork
nodes_vis <- node_graphdata %>%
  mutate(
    group = case_when(
      group %in% c("Song", "Album") ~ "Work",
      TRUE ~ group
    ),
    font_size = ifelse(group == "Genre", 20, 15),
    color = case_when(
      group == "Genre"  ~ "skyblue",
      group == "Work"   ~ "lightgreen",
      group == "Person" ~ "yellow",
      TRUE              ~ "gray"
    )
  ) %>%
  mutate(font = map(font_size, ~ list(size = .x)))

edges_vis <- edge_graphdata_final

# 📌 Step 11: interactive visualization
visNetwork(nodes = nodes_vis, edges = edges_vis, width = "100%", height = "700px") %>%
  visNodes(
    borderWidth = 2,
    borderWidthSelected = 4,
    shadow = TRUE,
    color = list(
      border = "black",
      highlight = list(border = "black", background = "orange")
    )
  ) %>%
  visEdges(arrows = "to", smooth = FALSE) %>%
  visOptions(
    highlightNearest = TRUE,
    nodesIdSelection = list(enabled = TRUE, useLabels = TRUE)
  ) %>%
  visInteraction(zoomView = TRUE, dragNodes = TRUE) %>%
  visLayout(randomSeed = 2025) %>%
  visGroups(groupname = "Genre",  color = "skyblue") %>%
  visGroups(groupname = "Work",   color = "lightgreen") %>%
  visGroups(groupname = "Person", color = "yellow") %>%
  visLegend(
  useGroups = TRUE,
  position = "right",
  main = "Node Type & Size Meaning"
)%>%
  visPhysics(
    solver = "forceAtlas2Based",
    forceAtlas2Based = list(gravitationalConstant = -50),
    stabilization = list(enabled = TRUE, iterations = 100)
  )

```

```{r}
library(dplyr)
library(ggplot2)
library(forcats)

library(ggplot2)
library(dplyr)
library(forcats)

# 1. 计算每个 person 的 in-degree
person_indegree_tbl <- work_to_person_edges %>%
  count(to, name = "in_degree") %>%
  left_join(node_person, by = c("to" = "id")) %>%
  select(id = to, name_person = label, in_degree)


# 2. 作图
ggplot(person_indegree_tbl, aes(x = fct_reorder(name_person, in_degree), y = in_degree)) +
  geom_col(fill = "#FFA500") +
  coord_flip() +
  labs(
    title = "Oceanus Folk Influence by Person",
    subtitle = "Number of Oceanus Folk-influenced works each artist participated in",
    x = "Artist",
    y = "Number of Influenced Works (In-Degree)"
  ) +
  theme_minimal(base_size = 13)


```

top10

```{r}
top10_tbl <- person_indegree_tbl %>%
  arrange(desc(in_degree)) %>%
  slice_head(n = 10)

```

```{r}
ggplot(top10_tbl, 
       aes(x = fct_reorder(name_person, in_degree, .desc = TRUE), y = in_degree)) +
  geom_col(fill = "#FFA500") +
  geom_text(aes(label = in_degree), hjust = -0.1, size = 4) +
  coord_flip() +
  labs(
    title = "Top 10 Most Influenced Artists by Oceanus Folk",
    subtitle = "Ranked by number of Oceanus Folk-influenced works participated in",
    x = "Artist",
    y = "Number of Influenced Works (In-Degree)"
  ) +
  theme_minimal(base_size = 13) +
  ylim(0, max(top10_tbl$in_degree) * 1.15)  # 给文本预留右边空间

```

--
