---
title: "Take Home Exercise 02"
author: "Zhang Jinghan"
date-modified: "last-modified"
format: html
execute:
  eval: true
  echo: true
  warning: false
  freeze: true #防止重复render
---

tidygraph: wrap of graph building code

```{r}
pacman::p_load(tidyverse, jsonlite,   #tidyverse全家桶里面把各种gg都包括了
               SmartEDA, tidygraph, 
               ggraph)
```

Inspect Structure

```{r}
# 读取 JSON 文件
kg <- fromJSON("data/MC1_graph.json")

# 查看结构
str(kg, max.level = 1)  #只看最高level的data structure

```

Extract and Inspect

```{r}
# 提取并检查节点和边数据
nodes_tbl <- as_tibble(kg$nodes)
edges_tbl <- as_tibble(kg$links)
```

Initial EDA

```{r}
ggplot(data = edges_tbl,
       aes(y=`Edge Type`))+
  geom_bar()
```

```{r}
ggplot(data = nodes_tbl,
       aes(y=`Node Type`))+
  geom_bar()
```

# creating knowledge graph

This is Step 1: mapping from node id to row index

```{r}
id_map <- tibble(id=nodes_tbl$id,
                 index = seq_len(
                   nrow(nodes_tbl)
                 ))
#处理后就没有id=0了
```

Step 2 :Map source and target IDs to row indices

```{r}
edges_tbl <- edges_tbl %>%
  left_join(id_map, by = c("source" = "id")) %>%
  rename(from = index) %>%
  left_join(id_map, by = c("target" = "id")) %>%
  rename(to = index)
```

Step 3: Filter out any unmatched

```{r}
edges_tbl<- edges_tbl%>%
  filter(!is.na(from), !is.na(to))
```

```{r}
edges_tbl <- edges_tbl %>%
  filter(!is.na(from), !is.na(to))

```

Step 4: Creating the graph

```{r}
graph <- tbl_graph(
  nodes = nodes_tbl,
  edges = edges_tbl,
  directed = kg$directed  # 从你图片中可见 directed 参数由 kg 控制
)

```

Visualising the knowledge graph

```{r}
set.seed(1234)
```

Visualising the whole graph

```{r}
#| eval: false
ggraph(graph, layout = "fr") +      #先plotlink再plot node 这样重叠顺序才对
  geom_edge_link(alpha = 0.3, 
                 colour = "gray") +
  geom_node_point(aes(color = `Node Type`), 
                  size = 1) +
  geom_node_text(aes(label = name), 
                 repel = TRUE, 
                 size = 2.5) +
  theme_void()

```

----------------------------
Step 1: Filter edges to only "MemberOf"
```{r}
graph_memberof<-graph %>%
  activate(edges)%>%
  filter(`Edge Type` == "MemberOf")
```

Step 2: Extract only connected nodes (i.e., use in these edges)
```{r}
used_node_indices <- graph_memberof %>%
  activate(edges) %>%
  as_tibble()%>%
  select(from, to)%>%
  unlist()%>%
  unique()

```

Step 3
```{r}
graph_memberof <- graph_memberof %>%
  activate(nodes) %>%
  mutate(row_id = row_number()) %>%
  filter(row_id %in% used_node_indices) %>%
  select(-row_id)  # optional cleanup

```


Plot the subplot1
```{r}
ggraph(graph_memberof, layout = "fr") +
  geom_edge_link(alpha = 0.5, 
                 colour = "gray") +
  geom_node_point(aes(color = `Node Type`), 
                  size = 1) +
  geom_node_text(aes(label = name), 
                 repel = TRUE, 
                 size = 2.5) +
  theme_void()
```

