---
title: "Hands-on Exercise 08"
author: "Zhang Jinghan"
date-modified: "last-modified"
format: html
execute:
  eval: true
  echo: true
  warning: false
  freeze: true #防止重复render
---

# **21  Choropleth Mapping with R**

## **21.1 Overview**

Choropleth mapping involves the symbolisation of enumeration units, such as countries, provinces, states, counties or census units, using area patterns or graduated colors. For example, a social scientist may need to use a choropleth map to portray the spatial distribution of aged population of Singapore by Master Plan 2014 Subzone Boundary.

In this chapter, you will learn how to plot functional and truthful choropleth maps by using an R package called [**tmap**](https://r-tmap.github.io/tmap/) package.

## **21.2 Getting Started**

In this hands-on exercise, the key R package use is [**tmap**](https://cran.r-project.org/web/packages/tmap/) package in R. Beside **tmap** package, four other R packages will be used. They are:

-   [**readr**](https://readr.tidyverse.org/) for importing delimited text file,

-   [**tidyr**](https://tidyr.tidyverse.org/) for tidying data,

-   [**dplyr**](https://dplyr.tidyverse.org/) for wrangling data and

-   [**sf**](https://cran.r-project.org/web/packages/sf/index.html) for handling geospatial data.

Among the four packages, **readr**, **tidyr** and **dplyr** are part of **tidyverse** package.

The code chunk below will be used to install and load these packages in RStudio.

```{r}
pacman::p_load(sf, tmap, tidyverse)
```

## **21.3 Importing Data into R**

### **21.3.1 The Data**

Two data set will be used to create the choropleth map. They are:

-   Master Plan 2014 Subzone Boundary (Web) (i.e. `MP14_SUBZONE_WEB_PL`) in ESRI shapefile format. It can be downloaded at [data.gov.sg](https://data.gov.sg/) This is a geospatial data. It consists of the geographical boundary of Singapore at the planning subzone level. The data is based on URA Master Plan 2014.

-   Singapore Residents by Planning Area / Subzone, Age Group, Sex and Type of Dwelling, June 2011-2020 in csv format (i.e. `respopagesextod2011to2020.csv`). This is an aspatial data fie. It can be downloaded at [Department of Statistics, Singapore](https://www.singstat.gov.sg/) Although it does not contain any coordinates values, but it’s PA and SZ fields can be used as unique identifiers to geocode to `MP14_SUBZONE_WEB_PL` shapefile.

### **21.3.2 Importing Geospatial Data into R**

The code chunk below uses the *st_read()* function of **sf** package to import `MP14_SUBZONE_WEB_PL` shapefile into R as a simple feature data frame called `mpsz`.

```{r}
mpsz <- st_read(dsn = "data/geospatial", 
                layer = "MP14_SUBZONE_WEB_PL")
```

You can examine the content of `mpsz` by using the code chunk below.

```{r}
mpsz
```

### **21.3.3 Importing Attribute Data into R**

Next, we will import *respopagsex2011to2020.csv* file into RStudio and save the file into an R dataframe called *popagsex*.

The task will be performed by using *read_csv()* function of **readr** package as shown in the code chunk below.

```{r}
popdata <- read_csv("data/aspatial/respopagesextod2011to2020.csv")
```

### **21.3.4 Data Preparation**

Before a thematic map can be prepared, you are required to prepare a data table with year 2020 values. The data table should include the variables PA, SZ, YOUNG, ECONOMY ACTIVE, AGED, TOTAL, DEPENDENCY.

-   YOUNG: age group 0 to 4 until age groyup 20 to 24,

-   ECONOMY ACTIVE: age group 25-29 until age group 60-64,

-   AGED: age group 65 and above,

-   TOTAL: all age group, and

-   DEPENDENCY: the ratio between young and aged against economy active group

#### 21.3.4.1 Data wrangling

The following data wrangling and transformation functions will be used:

-   *pivot_wider()* of **tidyr** package, and

-   *mutate()*, *filter()*, *group_by()* and *select()* of **dplyr** package

```{r}
popdata2020 <- popdata %>%
  filter(Time == 2020) %>%
  group_by(PA, SZ, AG) %>%
  summarise(`POP` = sum(`Pop`)) %>%
  ungroup() %>%
  pivot_wider(names_from=AG, 
              values_from=POP) %>%
  mutate(YOUNG = rowSums(.[3:6])
         +rowSums(.[12])) %>%
mutate(`ECONOMY ACTIVE` = rowSums(.[7:11])+
rowSums(.[13:15]))%>%
mutate(`AGED`=rowSums(.[16:21])) %>%
mutate(`TOTAL`=rowSums(.[3:21])) %>%  
mutate(`DEPENDENCY` = (`YOUNG` + `AGED`)
/`ECONOMY ACTIVE`) %>%
  select(`PA`, `SZ`, `YOUNG`, 
       `ECONOMY ACTIVE`, `AGED`, 
       `TOTAL`, `DEPENDENCY`)
```

#### 21.3.4.2 Joining the attribute data and geospatial data

Before we can perform the georelational join, one extra step is required to convert the values in PA and SZ fields to uppercase. This is because the values of *PA* and *SZ* fields are made up of upper- and lowercase. On the other, hand the *SUBZONE_N* and *PLN_AREA_N* are in uppercase.

```{r}
popdata2020 <- popdata2020 %>%
  mutate(across(c(PA, SZ), toupper)) %>%
  filter(`ECONOMY ACTIVE` > 0)
```

Next, *left_join()* of **dplyr** is used to join the geographical data and attribute table using planning subzone name e.g. *SUBZONE_N* and *SZ* as the common identifier.

```{r}
mpsz_pop2020 <- left_join(mpsz, popdata2020,
                          by = c("SUBZONE_N" = "SZ"))
```

Thing to learn from the code chunk above:

-   *left_join()* of **dplyr** package is used with `mpsz` simple feature data frame as the left data table is to ensure that the output will be a simple features data frame.

```{r}
write_rds(mpsz_pop2020, "data/rds/mpszpop2020.rds")
```

## **21.4 Choropleth Mapping Geospatial Data Using *tmap***

Two approaches can be used to prepare thematic map using *tmap*, they are:

-   Plotting a thematic map quickly by using *qtm()*.

-   Plotting highly customisable thematic map by using tmap elements.

### **21.4.1 Plotting a choropleth map quickly by using *qtm()***

The easiest and quickest to draw a choropleth map using **tmap** is using *qtm()*. It is concise and provides a good default visualisation in many cases.

The code chunk below will draw a cartographic standard choropleth map as shown below.\

```{r}
tmap_mode("plot")
qtm(mpsz_pop2020, 
    fill = "DEPENDENCY")
```

Things to learn from the code chunk above:

-   *tmap_mode()* with “plot” option is used to produce a static map. For interactive mode, “view” option should be used.

-   *fill* argument is used to map the attribute (i.e. DEPENDENCY)

### **21.4.2 Creating a choropleth map by using *tmap*’s elements**

Despite its usefulness of drawing a choropleth map quickly and easily, the disadvantge of *qtm()* is that it makes aesthetics of individual layers harder to control. To draw a high quality cartographic choropleth map as shown in the figure below, **tmap**’s drawing elements should be used.

```{r}
tm_shape(mpsz_pop2020)+
  tm_polygons(fill = "DEPENDENCY", 
              fill.scale = tm_scale_intervals(
                style = "quantile",
                n = 5,
                values = "brewer.blues"),
              fill.legend = tm_legend(
                title = "Dependency ratio")) +
  tm_title("Distribution of Dependency Ratio by planning subzone") +
  tm_layout(frame = TRUE) +
  tm_borders(fill_alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_grid(alpha =0.2) +
  tm_credits("Source: Planning Sub-zone boundary from Urban Redevelopment Authorithy (URA)\n and Population data from Department of Statistics DOS", 
             position = c("left", "bottom"))
```

In the following sub-section, we will share with you tmap functions that used to plot these elements.

#### 21.4.2.1 Drawing a base map

The basic building block of **tmap** is *tm_shape()* followed by one or more layer elemments such as *tm_fill()* and *tm_polygons()*.

In the code chunk below, *tm_shape()* is used to define the input data (i.e *mpsz_pop2020*) and *tm_polygons()* is used to draw the planning subzone polygons\

```{r}
tm_shape(mpsz_pop2020) +
  tm_polygons()
```

#### 21.4.2.2 Drawing a choropleth map using *tm_polygons()*

To draw a choropleth map showing the geographical distribution of a selected variable by planning subzone, we just need to assign the target variable such as *Dependency* to *tm_polygons()*.

```{r}
tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY")
```

Things to learn from *tm_polygons()*:

-   The default interval binning used to draw the choropleth map is called “pretty”. A detailed discussion of the data classification methods supported by **tmap** will be provided in sub-section 4.3.

-   The default colour scheme used is `YlOrRd` of ColorBrewer. You will learn more about the color scheme in sub-section 4.4.

-   By default, Missing value will be shaded in grey.

#### 21.4.2.3 Drawing a choropleth map using *tm_fill()* and \*tm_border()\*\*

Actually, *tm_polygons()* is a wraper of *tm_fill()* and *tm_border()*. *tm_fill()* shades the polygons by using the default colour scheme and *tm_borders()* adds the borders of the shapefile onto the choropleth map.

The code chunk below draws a choropleth map by using *tm_fill()* alone.

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY")
```

Notice that the planning subzones are shared according to the respective dependecy values

To add the boundary of the planning subzones, tm_borders will be used as shown in the code chunk below.

```{r}
tm_shape(mpsz_pop2020)+
  tm_polygons(fill = "DEPENDENCY") +
  tm_borders(lwd = 0.01,  
             fill_alpha = 0.1)
```

Notice that light-gray border lines have been added on the choropleth map.

The *alpha* argument is used to define transparency number between 0 (totally transparent) and 1 (not transparent). By default, the alpha value of the col is used (normally 1).

Beside *alpha* argument, there are three other arguments for *tm_borders()*, they are:

-   *col* = border colour,

-   *lwd* = border line width. The default is 1, and

-   *lty* = border line type. The default is “solid”.

### **21.4.3 Data classification methods of tmap**

Most choropleth maps employ some methods of data classification. The point of classification is to take a large number of observations and group them into data ranges or classes.

**tmap** provides a total ten data classification methods, namely: *fixed*, *sd*, *equal*, *pretty* (default), *quantile*, *kmeans*, *hclust*, *bclust*, *fisher*, and *jenks*.

To define a data classification method, the *style* argument of *tm_fill()* or *tm_polygons()* will be used.

#### 21.4.3.1 Plotting choropleth maps with built-in classification methods

The code chunk below shows a quantile data classification that used 5 classes.

```{r}
tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY",
      fill.scale = tm_scale_intervals(
        style = "jenks",
        n = 5)) +
  tm_borders(fill_alpha = 0.5)
```

In the code chunk below, *equal* data classification method is used.\

```{r}
tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY",
      fill.scale = tm_scale_intervals(
        style = "equal",
        n = 5)) +
  tm_borders(fill_alpha = 0.5)
```

Notice that the distribution of quantile data classification method are more evenly distributed then equal data classification method.

> **Warning: Maps Lie!**

> DIY: Using what you had learned, prepare choropleth maps by using different classification methods supported by tmap and compare their differences.

> DIY: Preparing choropleth maps by using similar classification method but with different numbers of classes (i.e. 2, 6, 10, 20). Compare the output maps, what observation can you draw?

#### 21.4.3.2 Plotting choropleth map with custome break

For all the built-in styles, the category breaks are computed internally. In order to override these defaults, the breakpoints can be set explicitly by means of the *breaks* argument to the *tm_fill()*. It is important to note that, in **tmap** the breaks include a minimum and maximum. As a result, in order to end up with n categories, n+1 elements must be specified in the *breaks* option (the values must be in increasing order).

Before we get started, it is always a good practice to get some descriptive statistics on the variable before setting the break points. Code chunk below will be used to compute and display the descriptive statistics of **DEPENDENCY** field.

```{r}
summary(mpsz_pop2020$DEPENDENCY)
```

With reference to the results above, we set break point at 0.60, 0.70, 0.80, and 0.90. In addition, we also need to include a minimum and maximum, which we set at 0 and 100. Our *breaks* vector is thus c(0, 0.60, 0.70, 0.80, 0.90, 1.00)

Now, we will plot the choropleth map by using the code chunk below.

```{r}
tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY",
          breaks = c(0, 0.60, 0.70, 0.80, 0.90, 1.00)) +
  tm_borders(fill_alpha = 0.5)
```

### **21.4.4 Colour Scheme**

**tmap** supports colour ramps either defined by the user or a set of predefined colour ramps from the **RColorBrewer** package.

#### 21.4.4.1 Using ColourBrewer palette

To change the colour, we assign the preferred colour to *values* argument of *tm_scale_intervals()* as shown in the code chunk below.

```{r}
tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY",
      fill.scale = tm_scale_intervals(
        style = "quantile",
        n = 5,
        values = "brewer.greens")) +
  tm_borders(fill_alpha = 0.5)
```

Notice that the choropleth map is shaded in green.

To reverse the colour shading, add a “-” prefix.

```{r}
tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY",
      fill.scale = tm_scale_intervals(
        style = "quantile",
        n = 5,
        values = "-brewer.greens")) +
  tm_borders(fill_alpha = 0.5)
```

Notice that the colour scheme has been reversed.

### **21.4.5 Map Layouts**

Map layout refers to the combination of all map elements into a cohensive map. Map elements include among others the objects to be mapped, the title, the scale bar, the compass, margins and aspects ratios. Colour settings and data classification methods covered in the previous section relate to the palette and break-points are used to affect how the map looks.

#### 21.4.5.1 Map Legend

In **tmap**, several `tm_legend()` options are provided to change the placement, format and appearance of the legend.

```{r}
tm_shape(mpsz_pop2020)+
  tm_polygons("DEPENDENCY",
      fill.scale = tm_scale_intervals(
        style = "jenks",
        n = 5,
        values = "brewer.greens"),
      fill.legend = tm_legend(
        title = "Dependency ratio")) +
  tm_borders(fill_alpha = 0.5) +
  tm_title("Distribution of Dependency Ratio by planning subzone \n(Jenks classification)")
```

#### 21.4.5.2 Map style

**tmap** allows a wide variety of layout settings to be changed. They can be called by using *tmap_style()*.

The code chunk below shows the *classic* style is used.

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "-Greens") +
  tm_borders(alpha = 0.5) +
  tmap_style("classic")
```

#### 21.4.5.3 Cartographic Furniture

Beside map style, **tmap** also also provides arguments to draw other map furniture such as compass, scale bar and grid lines.

In the code chunk below, `tm_compass()`, `tm_scale_bar()` and `tm_grid()` are used to add compass, scale bar and grid lines onto the choropleth map.

```{r}
tm_shape(mpsz_pop2020)+
  tm_polygons(fill = "DEPENDENCY", 
              fill.scale = tm_scale_intervals(
                style = "quantile",
                n = 5,
                values = "brewer.blues"),
              fill.legend = tm_legend(
                title = "Dependency ratio")) +
  tm_title("Distribution of Dependency Ratio by planning subzone") +
  tm_layout(frame = TRUE) +
  tm_borders(fill_alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_grid(alpha =0.2) +
  tm_credits("Source: Planning Sub-zone boundary from Urban Redevelopment Authorithy (URA)\n and Population data from Department of Statistics DOS", 
             position = c("left", "bottom"))
```

To reset the default style, refer to the code chunk below.

```{r}
tmap_style("white")
```

### **21.4.6 Drawing Small Multiple Choropleth Maps**

**Small multiple maps**, also referred to as **facet maps**, are composed of many maps arrange side-by-side, and sometimes stacked vertically. Small multiple maps enable the visualisation of how spatial relationships change with respect to another variable, such as time.

In **tmap**, small multiple maps can be plotted in three ways:

-   by assigning multiple values to at least one of the asthetic arguments,

-   by defining a group-by variable in *tm_facets()*, and

-   by creating multiple stand-alone maps with *tmap_arrange()*.

#### 21.4.6.1 By assigning multiple values to at least one of the aesthetic arguments

In this example, small multiple choropleth maps are created by defining ***ncols*** in **tm_fill()**

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill(c("YOUNG", "AGED"),
          style = "equal", 
          palette = "Blues") +
  tm_layout(legend.position = c("right", "bottom")) +
  tm_borders(alpha = 0.5) +
  tmap_style("white")
```

In this example, small multiple choropleth maps are created by assigning multiple values to at least one of the aesthetic arguments

```{r}
tm_shape(mpsz_pop2020)+ 
  tm_polygons(c("DEPENDENCY","AGED"),
          style = c("equal", "quantile"), 
          palette = list("Blues","Greens")) +
  tm_layout(legend.position = c("right", "bottom"))
```

#### 21.4.6.2 By defining a group-by variable in *tm_facets()*

In this example, multiple small choropleth maps are created by using `tm_facets()`.

```{r}
tm_shape(mpsz_pop2020) +
  tm_fill("DEPENDENCY",
          style = "quantile",
          palette = "Blues",
          thres.poly = 0) + 
  tm_facets(by="REGION_N", 
            free.coords=TRUE) +
  tm_layout(legend.show = FALSE,
            title.position = c("center", "center"), 
            title.size = 20) +
  tm_borders(alpha = 0.5)
```

#### 21.4.6.3 By creating multiple stand-alone maps with `tmap_arrange()`

In this example, multiple small choropleth maps are created by creating multiple stand-alone maps with `tmap_arrange()`.

```{r}
youngmap <- tm_shape(mpsz_pop2020)+ 
  tm_polygons("YOUNG", 
              style = "quantile", 
              palette = "Blues")

agedmap <- tm_shape(mpsz_pop2020)+ 
  tm_polygons("AGED", 
              style = "quantile", 
              palette = "Blues")

tmap_arrange(youngmap, agedmap, asp=1, ncol=2)
```

### **21.4.7 Mappping Spatial Object Meeting a Selection Criterion**

Instead of creating small multiple choropleth map, you can also use selection funtion to map spatial objects meeting the selection criterion.

```{r}
tm_shape(mpsz_pop2020[mpsz_pop2020$REGION_N=="CENTRAL REGION", ])+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "Blues", 
          legend.hist = TRUE, 
          legend.is.portrait = TRUE,
          legend.hist.z = 0.1) +
  tm_layout(legend.outside = TRUE,
            legend.height = 0.45, 
            legend.width = 5.0,
            legend.position = c("right", "bottom"),
            frame = FALSE) +
  tm_borders(alpha = 0.5)
```

# **22  Visualising Geospatial Point Data**

## **Overview**

Proportional symbol maps (also known as graduate symbol maps) are a class of maps that use the visual variable of size to represent differences in the magnitude of a discrete, abruptly changing phenomenon, e.g. counts of people. Like choropleth maps, you can create classed or unclassed versions of these maps. The classed ones are known as range-graded or graduated symbols, and the unclassed are called proportional symbols, where the area of the symbols are proportional to the values of the attribute being mapped. In this hands-on exercise, you will learn how to create a proportional symbol map showing the number of wins by Singapore Pools’ outlets using an R package called **tmap**.

## **Learning outcome**

By the end of this hands-on exercise, you will acquire the following skills by using appropriate R packages:

-   To import an aspatial data file into R.

-   To convert it into simple point feature data frame and at the same time, to assign an appropriate projection reference to the newly create simple point feature data frame.

-   To plot interactive proportional symbol maps.

## **Getting Started**

Before we get started, we need to ensure that **tmap** package of R and other related R packages have been installed and loaded into R.

```{r}
pacman::p_load(sf, tmap, tidyverse)
```

## **Geospatial Data Wrangling**

## **The data**

The data set use for this hands-on exercise is called *SGPools_svy21*. The data is in csv file format.

Figure below shows the first 15 records of SGPools_svy21.csv. It consists of seven columns. The XCOORD and YCOORD columns are the x-coordinates and y-coordinates of SingPools outlets and branches. They are in [Singapore SVY21 Projected Coordinates System](https://www.sla.gov.sg/sirent/CoordinateSystems.aspx).

## **Data Import and Preparation**

The code chunk below uses *read_csv()* function of **readr** package to import *SGPools_svy21.csv* into R as a tibble data frame called *sgpools*.

```{r}
sgpools <- read_csv("data/aspatial/SGPools_svy21.csv")
```

After importing the data file into R, it is important for us to examine if the data file has been imported correctly.

The code chunk below shows list() is used to do the job.

```{r}
list(sgpools) 
```

## **Creating a sf data frame from an aspatial data frame**

The code chunk below converts sgpools data frame into a simple feature data frame by using *st_as_sf()* of **sf** packages

```{r}
sgpools_sf <- st_as_sf(sgpools, 
                       coords = c("XCOORD", "YCOORD"),
                       crs= 3414)
```

Things to learn from the arguments above:

-   The *coords* argument requires you to provide the column name of the x-coordinates first then followed by the column name of the y-coordinates.

-   The *crs* argument required you to provide the coordinates system in epsg format. [EPSG: 3414](https://epsg.io/3414) is Singapore SVY21 Projected Coordinate System. You can search for other country’s epsg code by refering to [epsg.io](https://epsg.io/).

Figure below shows the data table of *sgpools_sf*. Notice that a new column called geometry has been added into the data frame.

You can display the basic information of the newly created sgpools_sf by using the code chunk below.

```{r}
list(sgpools_sf)
```

The output shows that sgppols_sf is in point feature class. It’s epsg ID is 3414. The bbox provides information of the extend of the geospatial data.

## **Drawing Proportional Symbol Map**

To create an interactive proportional symbol map in R, the view mode of tmap will be used.

The code churn below will turn on the interactive mode of tmap.

```{r}
tmap_mode("view")
```

## **It all started with an interactive point symbol map**

The code chunks below are used to create an interactive point symbol map.

```{r}
tm_shape(sgpools_sf) + 
  tm_bubbles(fill = "red",
           size = 1,
           col = "black",
           lwd = 1)

```

## **Lets make it proportional**

To draw a proportional symbol map, we need to assign a numerical variable to the size visual attribute. The code chunks below show that the variable *Gp1Gp2Winnings* is assigned to size visual attribute.

```{r}
tm_shape(sgpools_sf) + 
  tm_bubbles(fill = "red",
             size = "Gp1Gp2 Winnings",
             col = "black",
             lwd = 1)
```

## **Lets give it a different colour**

The proportional symbol map can be further improved by using the colour visual attribute. In the code chunks below, *OUTLET_TYPE* variable is used as the colour attribute variable.

```{r}
tm_shape(sgpools_sf) + 
  tm_bubbles(fill = "OUTLET TYPE", 
             size = "Gp1Gp2 Winnings",
             col = "black",
             lwd = 1)
```

## **I have a twin brothers :)**

An impressive and little-know feature of **tmap**’s view mode is that it also works with faceted plots. The argument *sync* in *tm_facets()* can be used in this case to produce multiple maps with synchronised zoom and pan settings.

```{r}
tm_shape(sgpools_sf) + 
  tm_bubbles(fill = "OUTLET TYPE", 
             size = "Gp1Gp2 Winnings",
             col = "black",
             lwd = 1) + 
  tm_facets(by= "OUTLET TYPE",
            nrow = 1,
            sync = TRUE)
```

Before you end the session, it is wiser to switch **tmap**’s Viewer back to plot mode by using the code chunk below.

```{r}
tmap_mode("plot")
```

# **23  Analytical Mapping**

## **23.1 Overview**

### **23.1.1 Objectives**

In this in-class exercise, you will gain hands-on experience on using appropriate R methods to plot analytical maps.

### **23.1.2 Learning outcome**

By the end of this in-class exercise, you will be able to use appropriate functions of tmap and tidyverse to perform the following tasks:

-   Importing geospatial data in rds format into R environment.

-   Creating cartographic quality choropleth maps by using appropriate tmap functions.

-   Creating rate map

-   Creating percentile map

-   Creating boxmap

## **23.2 Getting Started**

### **23.2.1 Installing and loading packages**

```{r}
pacman::p_load(tmap,tidyverse,sf)
```

### **23.2.2 Importing data**

For the purpose of this hands-on exercise, a prepared data set called *NGA_wp.rds* will be used. The data set is a polygon feature data.frame providing information on water point of Nigeria at the LGA level. You can find the data set in the *rds* sub-direct of the hands-on *data* folder.

```{r}
NGA_wp <- read_rds("data/rds/NGA_wp.rds")
```

## **23.3 Basic Choropleth Mapping**

### **23.3.1 Visualising distribution of non-functional water point**

```{r}
p1 <- tm_shape(NGA_wp) +
  tm_polygons(fill = "wp_functional",
             fill.scale = tm_scale_intervals(
               style = "equal",
               n = 10,
               values = "brewer.blues"),
             fill.legend = tm_legend(
               position = c("right", "bottom"))) +
  tm_borders(lwd = 0.1,
             fill_alpha = 1) +
  tm_title("Distribution of functional water point by LGAs")
```

```{r}
p2 <- tm_shape(NGA_wp) + 
  tm_polygons(fill = "total_wp", 
              fill.scale = tm_scale_intervals(
                style = "equal",
                n = 10,
                values = "brewer.blues"),
              fill.legend = tm_legend(
                position = c("right", "bottom"))) +
  tm_borders(lwd = 0.1, 
             fill_alpha = 1) + 
  tm_title("Distribution of total  water point by LGAs")
```

```{r}
tmap_arrange(p2, p1, nrow = 1)
```

## **23.4 Choropleth Map for Rates**

In much of our readings we have now seen the importance to map rates rather than counts of things, and that is for the simple reason that water points are not equally distributed in space. That means that if we do not account for how many water points are somewhere, we end up mapping total water point size rather than our topic of interest.

### **23.4.1 Deriving Proportion of Functional Water Points and Non-Functional Water Points**

We will tabulate the proportion of functional water points and the proportion of non-functional water points in each LGA. In the following code chunk, `mutate(`) from **dplyr** package is used to derive two fields, namely *pct_functional* and *pct_nonfunctional*.

```{r}
NGA_wp <- NGA_wp %>%
  mutate(pct_functional = wp_functional/total_wp) %>%
  mutate(pct_nonfunctional = wp_nonfunctional/total_wp)
```

### **23.4.2 Plotting map of rate**

```{r}
tm_shape(NGA_wp) +
  tm_polygons("pct_functional",
              fill.scale = tm_scale_intervals(
                style = "equal",
                n = 10,
                values = "brewer.blues"),
              fill.legend = tm_legend(
                position = c("right", "bottom"))) + 
  tm_borders(lwd = 0.1,
             fill_alpha = 1) +
  tm_title("Rate map of functional water point by LGAs")
```

## **23.5 Extreme Value Maps**

Extreme value maps are variations of common choropleth maps where the classification is designed to highlight extreme values at the lower and upper end of the scale, with the goal of identifying outliers. These maps were developed in the spirit of spatializing EDA, i.e., adding spatial features to commonly used approaches in non-spatial EDA (Anselin 1994).

### **23.5.1 Percentile Map**

The percentile map is a special type of quantile map with six specific categories: 0-1%,1-10%, 10-50%,50-90%,90-99%, and 99-100%. The corresponding breakpoints can be derived by means of the base R quantile command, passing an explicit vector of cumulative probabilities as c(0,.01,.1,.5,.9,.99,1). Note that the begin and endpoint need to be included.

#### 23.5.1.1 Data Preparation

Step 1: Exclude records with NA by using the code chunk below.

```{r}
NGA_wp <- NGA_wp %>%
  drop_na()
```

Step 2: Creating customised classification and extracting values

```{r}
percent <- c(0,.01,.1,.5,.9,.99,1)
var <- NGA_wp["pct_functional"] %>%
  st_set_geometry(NULL)
quantile(var[,1], percent)
```

::: callout-important
When variables are extracted from an sf data.frame, the geometry is extracted as well. For mapping and spatial manipulation, this is the expected behavior, but many base R functions cannot deal with the geometry. Specifically, the `quantile()` gives an error. As a result `st_set_geomtry(NULL)` is used to drop geomtry field.
:::

#### 23.5.1.2 Why writing functions?

Writing a function has three big advantages over using copy-and-paste:

-   You can give a function an evocative name that makes your code easier to understand.

-   As requirements change, you only need to update code in one place, instead of many.

-   You eliminate the chance of making incidental mistakes when you copy and paste (i.e. updating a variable name in one place, but not in another).

Source: [Chapter 19: Functions](https://r4ds.had.co.nz/functions.html#functions) of **R for Data Science**.

#### 23.5.1.3 Creating the get.var function

Firstly, we will write an R function as shown below to extract a variable (i.e. *wp_nonfunctional*) as a vector out of an sf data.frame.

-   arguments:

    -   vname: variable name (as character, in quotes)

    -   df: name of sf data frame

-   returns:

    -   v: vector with values (without a column name)

```{r}
get.var <- function(vname,df) {
  v <- df[vname] %>% 
    st_set_geometry(NULL)
  v <- unname(v[,1])
  return(v)
}
```

#### 23.5.1.4 A percentile mapping function

Next, we will write a percentile mapping function by using the code chunk below.

```{r}
percentmap <- function(vnam, df, legtitle=NA, mtitle="Percentile Map"){
  percent <- c(0,.01,.1,.5,.9,.99,1)
  var <- get.var(vnam, df)
  bperc <- quantile(var, percent)
  tm_shape(df) +
  tm_polygons() +
  tm_shape(df) +
     tm_polygons(vnam,
             title=legtitle,
             breaks=bperc,
             palette="Blues",
          labels=c("< 1%", "1% - 10%", "10% - 50%", "50% - 90%", "90% - 99%", "> 99%"))  +
  tm_borders() +
  tm_layout(main.title = mtitle, 
            title.position = c("right","bottom"))
}
```

#### 23.5.1.5 Test drive the percentile mapping function

To run the function, type the code chunk as shown below.

```{r}
percentmap("total_wp", NGA_wp)
```

Note that this is just a bare bones implementation. Additional arguments such as the title, legend positioning just to name a few of them, could be passed to customise various features of the map.

### **23.5.2 Box map**

In essence, a box map is an augmented quartile map, with an additional lower and upper category. When there are lower outliers, then the starting point for the breaks is the minimum value, and the second break is the lower fence. In contrast, when there are no lower outliers, then the starting point for the breaks will be the lower fence, and the second break is the minimum value (there will be no observations that fall in the interval between the lower fence and the minimum value).

```{r}
ggplot(data = NGA_wp,
       aes(x = "",
           y = wp_nonfunctional)) +
  geom_boxplot()
```

-   Displaying summary statistics on a choropleth map by using the basic principles of boxplot.

-   To create a box map, a custom breaks specification will be used. However, there is a complication. The break points for the box map vary depending on whether lower or upper outliers are present.

#### 23.5.2.1 Creating the boxbreaks function

The code chunk below is an R function that creating break points for a box map.

-   arguments:

    -   v: vector with observations

    -   mult: multiplier for IQR (default 1.5)

-   returns:

    -   bb: vector with 7 break points compute quartile and fences

```{r}
boxbreaks <- function(v,mult=1.5) {
  qv <- unname(quantile(v))
  iqr <- qv[4] - qv[2]
  upfence <- qv[4] + mult * iqr
  lofence <- qv[2] - mult * iqr
  # initialize break points vector
  bb <- vector(mode="numeric",length=7)
  # logic for lower and upper fences
  if (lofence < qv[1]) {  # no lower outliers
    bb[1] <- lofence
    bb[2] <- floor(qv[1])
  } else {
    bb[2] <- lofence
    bb[1] <- qv[1]
  }
  if (upfence > qv[5]) { # no upper outliers
    bb[7] <- upfence
    bb[6] <- ceiling(qv[5])
  } else {
    bb[6] <- upfence
    bb[7] <- qv[5]
  }
  bb[3:5] <- qv[2:4]
  return(bb)
}
```

#### 23.5.2.2 Creating the get.var function

The code chunk below is an R function to extract a variable as a vector out of an sf data frame.

-   arguments:

    -   vname: variable name (as character, in quotes)

    -   df: name of sf data frame

-   returns:

    -   v: vector with values (without a column name)

```{r}
get.var <- function(vname,df) {
  v <- df[vname] %>% st_set_geometry(NULL)
  v <- unname(v[,1])
  return(v)
}
```

#### 23.5.2.3 Test drive the newly created function

Let’s test the newly created function

```{r}
var <- get.var("wp_nonfunctional", NGA_wp) 
boxbreaks(var)
```

#### 23.5.2.4 Boxmap function

The code chunk below is an R function to create a box map. - arguments: - vnam: variable name (as character, in quotes) - df: simple features polygon layer - legtitle: legend title - mtitle: map title - mult: multiplier for IQR - returns: - a tmap-element (plots a map)

```{r}
boxmap <- function(vnam, df, 
                   legtitle=NA,
                   mtitle="Box Map",
                   mult=1.5){
  var <- get.var(vnam,df)
  bb <- boxbreaks(var)
  tm_shape(df) +
    tm_polygons() +
  tm_shape(df) +
     tm_fill(vnam,title=legtitle,
             breaks=bb,
             palette="Blues",
          labels = c("lower outlier", 
                     "< 25%", 
                     "25% - 50%", 
                     "50% - 75%",
                     "> 75%", 
                     "upper outlier"))  +
  tm_borders() +
  tm_layout(main.title = mtitle, 
            title.position = c("left",
                               "top"))
}
```

```{r}
tmap_mode("plot")
boxmap("wp_nonfunctional", NGA_wp)
```
